---
title: "Youth NEET rates by Continent"
author: "Sahil Ali"
date: "2025-12-07"
output: html_document
---

This R markdown file will show the code for Youth NEET rates by continent

```{r}
countries <- read.csv("data sets/continents-according-to-our-world-in-data.csv",
                      stringsAsFactors = F)
youth_unemployment <- read.csv("data sets/youth-not-in-education-employment-training.csv",
                               stringsAsFactors = F)
```

Calculating the mean youth unemployment rate by Continent
Join with youth unemployment data 

```{r}

library(dplyr)

youth_unemployment2 <- countries %>%
  left_join(youth_unemployment, by = "Entity")
```

Removing and renaming certain columns, filtering out countries with no data

```{r}

youth_unemployment2 <- youth_unemployment2 %>%
  select(-Year.x, -Code.x) %>%
  rename(Country = Entity) %>%
  rename(Year = Year.y) %>%
  rename(Code = Code.y) %>%
  rename(y_unemployment = Share.of.youth.not.in.education..employment.or.training..total....of.youth.population.) %>%
  filter(!is.na(y_unemployment))
```

Calculate the change in youth unemployment every year

```{r}

youth_unemployment_change <- youth_unemployment2 %>%
  filter(Year >= 2015) %>%
  arrange(Country, Year) %>%              # Make sure order is correct
  group_by(Country) %>%                   # Calculate within each country
  mutate(pct_change = ((y_unemployment - lag(y_unemployment)) / lag(y_unemployment)) * 100) %>%
  ungroup()
```

Calculate the average youth unemployment change each year per continent

```{r}

continent_avg <- youth_unemployment_change %>%
  group_by(Continent, Year) %>%
  summarise(avg_pct_change = mean(pct_change, na.rm = TRUE), .groups = "drop")
```

Now to plot the graph

```{r}

library(ggplot2)

ggplot(continent_avg, aes(x = Year, y = avg_pct_change)) +
  geom_line(size = 1, colour = "steelblue") +
  geom_point(size = 1) +
  geom_hline(yintercept = 0, linetype = "dashed") +
  facet_wrap(~Continent) +
  labs(
    title = "Average % Change in Youth NEET rates by Continent (from 2015)",
    x = "Year",
    y = "Average % Change") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    axis.title = element_text(size = 10),
    legend.title = element_text(size = 14))

ggsave("Youth NEET Rates Change Plot.png", width = 10, height = 6, dpi = 300)
```


